---
title: 'Assignment 3'
author: "Alia Ajina"
date: "March 4, 2023"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: journal
    code_folding: show
    smooth_scroll: yes
    collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE)

library(tidyverse)
library(purrr)
library(ggpubr)
library(kableExtra)
```

## Looping assignment
Based on Naomi Tague's Rmarkdown "Looping"

Using our **autpower** function to estimate a distribution of power using pmap and for loops.

### Our highway

4 car types 

We have 4 different car types - and we know how often each occurs:

* car A  mass 20000 kg, area 22 m^2^
* car B mass 65000 kg, area 30 m^2^
* car C mass 38000 kg area 22 m^2^
* car D mass 5000 kg area 10 m^2^

Mean highway speed is 80 km/hr

SD of 10 km/hr


```{r sampling, eval=TRUE, echo=TRUE}
source("../R/autopower.R")

# Generating a dataframe to store information on our possible cars
possible_cars = data.frame(name = c("A", "B", "C", "D"),
                           mass=c(10000, 65000, 38000, 5000), 
                           area = c(22, 30, 22, 10))

# Convert our base speed from km/hr to m/s - in the future I will use a function for this
speed_base = 80 * 0.277

# Since there is 1 mass and 1 area for each car, using our function **autopower** we can estimate power for each car
possible_cars$power = autopower(V=speed_base, 
                                A = possible_cars$area, 
                                m=possible_cars$mass)

# Quick visualization of our powers per each car's area and mass                        
ggplot(possible_cars, aes(x=mass, y=power, fill=as.factor(area))) + 
  geom_col() + 
  labs(y="Power W", x="Mass (kg)", fill="Area"~(m^2))

```

**Figure 1** Quick visualization of power consumption associated with each car's area (m^2) and mass (kg).

### Building a highway

We will use *sample* here to generate speeds for 100 cars and look at two different ways to
repeat power calculation for different cars

We will then use *pmap* and *for*: to loop our results
* *pmap* for repetition - a efficient way of looping in R
* *for* for repetition - a more standard way of looping available in many langugaes

## Method 1: *pmap*

First, we'll use pmap to automate this task using a sample of 200 speeds with a log normal distribution and a mean of 80km/hr. The standard deviation (10km/hr) is 8% of the mean speed, 80 km/hr.


```{r withpmap}

possible_cars

# Creating our sample speeds and converting from km/hr to m/s
nsample = 200
mean_speed = log(80*0.277)

speeds = rlnorm(mean=mean_speed, sd=.08*mean_speed, nsample)
summary(speeds)

# Quick visualization of our distribution of speeds based on this random sample
# plot(density(speeds), ylab="Distribution of Speeds in (m/s)")

# Running each car for all speeds using pmap: runs a function for each value in a list of parameters, with other parameters set for each iteration

powerall1 = pmap(list(A = possible_cars$area, m=possible_cars$mass), autopower, V=speeds)

# Creating dataframe
powerall1 = as.data.frame(powerall1, col.names=possible_cars$name)


# Turning columns into rows to plot and table easier using pivot_longer
powerallr1 = powerall1 %>% pivot_longer(cols=everything(), 
                                        names_to="car", 
                                        values_to="power")

# Find mean and max powers and create table
means_pmap <- powerallr1 %>% 
  group_by(car) %>% 
  summarize(mean = mean(power),
            max = max(power))

means_pmap %>% 
  kable(caption = "Table 1: Sample statistics for car powers using pmap method",
        col.names = c("Car", "Mean power (m/s)", "Max power (m/s)")) %>% 
  kable_styling(full_width = F,
                bootstrap_options = "bordered",
                position = "center")

# Visualization of boxplots for method 1 - pmap

method1_plot = ggplot(powerallr1, aes(car, power, fill=car)) + 
  geom_boxplot() + 
  ggtitle("Method 1: Pmap") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(y = "Car", x = "Power") + theme_minimal()

method1_plot

```

**Figure 2** Boxplot displaying power consumption for each car using *pmap*.


## Method 2: Classic looping *for*

*pmap* works quickly but it is unique to R, so we will use *for* loops to compute our powers for each car as this is more common in other programming languages.

Again, this task uses a sample of 200 speeds with a log normal distribution and a mean of 80km/hr. The standard deviation (10km/hr) is 8% of the mean speed, 80 km/hr.

```{r withforloop}

possible_cars

# Creating our sample speeds and converting from km/hr to m/s
nsample = 200
mean_speed = log(80*0.277)

speeds = rlnorm(mean=mean_speed, sd=.08*mean_speed, nsample)
summary(speeds)

# Quick visualization of our distribution of speeds based on this random sample
# plot(density(speeds), ylab="Distribution of Speeds in (m/s)")

# Running each car for all speeds using a for loop that cycles through
# Creating dataframe
powerall2 = as.data.frame(matrix(nrow=length(speeds), 
                                 ncol=nrow(possible_cars)))


# Running *for* loop using our data from possible_cars
for (i in 1:nrow(possible_cars)) {
  powerall2[,i] = autopower(A=possible_cars$area[i], 
                            m=possible_cars$mass[i], 
                            V=speeds)
}

# Add column names
colnames(powerall2)=possible_cars$name


# Turning columns into rows for plotting and table
powerallr2 = powerall2 %>% pivot_longer(cols=everything(), 
                                        names_to="car", 
                                        values_to="power")

# Find mean and max powers and create table
means_pmap <- powerallr2 %>% 
  group_by(car) %>% 
  summarize(mean = mean(power),
            max = max(power))

means_pmap %>% 
  kable(caption = "Table 2: Sample statistics for car powers using for loop method",
        col.names = c("Car", "Mean power (m/s)", "Max power (m/s)")) %>% 
  kable_styling(full_width = F,
                bootstrap_options = "bordered",
                position = "center")

# Visualization of boxplots for method 2 - for loop
method2_plot = ggplot(powerallr2, aes(car, power, fill=car)) + 
  geom_boxplot() + 
  ggtitle("Method 2: For Loop") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(y = "Car", x = "Power") + theme_minimal()

method2_plot

```

**Figure 3** Boxplot displaying power consumption for each car using *for* loop.


## Both Methods Comparison

```{r boxplots combined}

# Combining the two boxplots (pmap and for loop methods) for comparison
ggarrange(method1_plot, method2_plot, nrow=1)

```

**Figure 4** Comparison of method 1 and method 2. As we used a random number generator for our samples, the power averages and ranges are different, however, the trends are the same in each scenario between car A, B, C, and D.
